\chapter{Hardware}

The Wokwi simulation described in \texttt{diagram.json} uses an Arduino Uno, a servo, buttons, LEDs, sensors, and a 16x2 LCD. The wiring matches the pins defined in the firmware, so the simulation faithfully represents the behavior expected on real hardware.

\section{Components}
\begin{itemize}
  \item Arduino Uno
  \item Servo for the turnstile
  \item 2 buttons (IN/OUT) with pull-down resistors
  \item Green and red LEDs (traffic light)
  \item Blue and orange LEDs (cooling/heating)
  \item White LED (humidity)
  \item Photoresistor (LDR)
  \item NTC sensor for temperature
  \item DHT22 sensor for humidity
  \item 4 yellow LEDs for the ceiling light (driven in parallel)
  \item 16x2 LCD
\end{itemize}

\section{Selection criteria}
Components were chosen for easy integration and the availability of stable libraries. The NTC provides a linear analog measure for the thermostat, while the DHT22 provides humidity with a simple digital interface. The 16x2 LCD balances cost and readability, with a standardized interface via \texttt{LiquidCrystal}.

\section{Pin map}
\begin{longtable}{@{}lll@{}}
\toprule
Function & Pin & Notes \\
\midrule
Turnstile servo & D3 & PWM \\
IN button & D6 & digital input \\
OUT button & D7 & digital input \\
Traffic light green LED & D4 & digital output \\
Traffic light red LED & D5 & digital output \\
Cooling LED & D8 & digital output \\
Heating LED & D9 & digital output \\
Ceiling light (PWM) & D10 & lighting dimmer \\
Humidity LED & D11 & digital output \\
Temperature sensor (NTC) & A0 & analog input \\
Photoresistor & A1 & analog input \\
LCD RS & D12 & \\
LCD EN & D13 & \\
LCD D4 & A2 & \\
LCD D5 & A3 & \\
LCD D6 & A4 & \\
LCD D7 & A5 & \\
DHT22 & D2 & pin defined in \texttt{humidity.h} \\
\bottomrule
\end{longtable}

\section{Wiring notes}
\begin{itemize}
  \item LEDs are connected with 220\,\textohm resistors to limit current.
  \item Buttons use 10\,k\textohm pull-down resistors to avoid floating readings.
  \item The DHT22 uses digital pin 2 and requires 5V power.
  \item The ceiling light is driven in PWM on D10, which simplifies intensity control.
\end{itemize}
