\chapter{Appendix: public APIs}

This appendix summarizes the main classes and functions exposed by the modules. APIs are designed to be minimal and easy to test, with a common \texttt{begin()} + \texttt{update()} pattern for control modules.

\section{Turnstile}
Manages people counting and servo actuation based on button inputs.
\begin{itemize}
  \item \texttt{Turnstile(int inButton, int outButton, int maxPeople)}
  \item \texttt{void begin(Servo* servo)}
  \item \texttt{void update()}
  \item \texttt{int getPeopleCount() const}
\end{itemize}

\section{Stoplight}
Translates capacity into a red/green visual signal.
\begin{itemize}
  \item \texttt{Stoplight(int greenPin, int redPin)}
  \item \texttt{void begin()}
  \item \texttt{void update(int currentPeople, int maxPeople)}
\end{itemize}

\section{Thermostat}
Controls heating and cooling with tolerance and an anti-toggle pause.
\begin{itemize}
  \item \texttt{Thermostat(int sensorPin, float desiredTemperature, int}\\
        \texttt{heatingPin, int coolingPin)}
  \item \texttt{void begin()}
  \item \texttt{bool update()}
  \item \texttt{void setTargetTemperature(float newTarget)}
  \item \texttt{float getTargetTemperature() const}
  \item \texttt{float getCurrentTemperature() const}
\end{itemize}

\section{HumidifierControl}
Controls humidity with sentinel handling in case of error.
\begin{itemize}
  \item \texttt{HumidifierControl(int humidifierPin, float targetHumidity)}
  \item \texttt{void begin()}
  \item \texttt{bool update()}
  \item \texttt{void setTargetHumidity(float newHumidityTarget)}
  \item \texttt{float getTargetHumidity() const}
  \item \texttt{float getCurrentHumidity() const}
\end{itemize}

\section{LightingControl}
Regulates lighting based on a target lux via PWM.
\begin{itemize}
  \item \texttt{LightingControl(int sensorPin, int dimmerPin, int targetLux)}
  \item \texttt{void begin()}
  \item \texttt{bool update()}
  \item \texttt{void setTargetLux(int newLuxTarget)}
  \item \texttt{int getTargetLux() const}
  \item \texttt{float getCurrentLux() const}
\end{itemize}

\section{DisplayPanel}
Manages the LCD UI with rotating pages and error messages.
\begin{itemize}
  \item \texttt{DisplayPanel(int rs, int en, int d4, int d5, int d6,}\\
        \texttt{int d7, int maxPeople)}
  \item \texttt{void begin()}
  \item \texttt{void update(int currentPeople, Thermostat\&,}\\
        \texttt{HumidifierControl\&, LightingControl\&)}
\end{itemize}

\section{Utility functions}
Reusable primitives for I/O and sensor readings.
\begin{itemize}
  \item \texttt{bool led(int pin, bool state)}
  \item \texttt{bool ledDimming(int pin, int pwm)}
  \item \texttt{float readTemperatureCelsius(int sensorPin)}
  \item \texttt{float readHumidity()}
  \item \texttt{float readLux(int light\_sensor\_pin)}
  \item \texttt{bool antiSufferingServo(int angle, Servo\& s)}
\end{itemize}
